### 2025-08-16 mixed-models.qmd
20250816 - tweak

Mixed modeling is a more flexible modeling framework than [analysis of variance](#sec-anova) (ANOVA).
Unlike [ANOVA](#sec-anova), mixed modeling can accommodate both continuous and categorical [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
[ANOVA](#sec-anova), cannot handle continuous [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
In addition, [ANOVA](#sec-anova) yields biased inferences when the same participants are exposed to the same items or stimuli [@Brauer2018].
Additionally, mixed models better handle missing data compared to [ANOVA](#sec-anova) [@Brauer2018].

>>>>

Mixed modeling is a more flexible modeling framework than [analysis of variance](#sec-anova) (ANOVA).
Unlike [ANOVA](#sec-anova), mixed modeling can accommodate both continuous and categorical [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
[ANOVA](#sec-anova), cannot handle continuous [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
In addition, [ANOVA](#sec-anova) yields biased inferences (i.e., a higher Type I error rate) when the same participants are exposed to multiple items or stimuli [i.e., responses by the same respondent tend to be correlated across items, and responses to the same item tend to be correlated across respondents; @Brauer2018].
Additionally, mixed models better handle missing data compared to [ANOVA](#sec-anova) [@Brauer2018].
